### 2020-06-23 sections/introduction.tex
makes changes suggested by SK

Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}, examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}). The Hubble classification scheme \citep{1926ApJ....64..321H}, and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.

@@

A majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and it is possible that their passage triggers star formation \citep{2013A&A...560A..59C}, though their main role may instead be to sweep up material, promoting the growth of Giant Molecular Clouds. \citep{2014IAUS..298..221D}. This rearrangement of disc gas can lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g. \citealt{2004ARA&A..42..603K}). Studies of spiral morphology have found interesting correlations between spiral morphology and other galactic properties, such as a correlation between spiral tightness and central mass concentration (\citealt{2019ApJ...871..194Y}, though \citealt{2017MNRAS.472.2263H} found no such relation) and tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure). These predictions and observations provide compelling reasons for investigating their underlying rules and dynamics, as doing so is essential for understanding the secular evolution of disc galaxies.

Our current understanding of the mechanisms which drive spiral growth and evolution suggests that each of the different forms of spiral galaxy may be triggered primarily by different processes. Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S}, \citealt{2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R}, \citealt{2009MNRAS.394...67A}, \citealt{2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and recurrent in nature \citep{1966ApJ...146..810J}.

One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}. The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most disc galaxies show spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.

Many simulations demonstrate that spirals do not maintain a constant pitch angle, and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}. Recent research suggests that spirals arms are transient in nature, and continually dissipate and re-form \citep{2014PASA...31...35D}. These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e. WASER, \citealt{1976ApJ...205..363M}, swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave. The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle, before dissipating to be reformed.

In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissipation. \citet{2019arXiv190910291P} propose a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time. They found that the distribution of pitch angles of their sample of 86 galaxies was consistent with this prediction, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.

We aim to test this idea of spiral winding using data from the \textit{Galaxy Builder} citizen science project for the spiral galaxies present in \citet{2020arXiv200610450L}. We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters produced by \textit{Galaxy Builder}. This methodology allows us to quantify the differences in pitch angles between arms in a single galaxy, as well as investigate the distribution of pitch angles in the galaxy population and investigate relationships between pitch angle and galaxy morphology.

Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of a stellar bar; simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}. Manifold theory is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds. One of the primary factors influencing the shape of this invariant manifold is the relative strength of the non-axisymmetric forcing caused by the bar, with stronger bars resulting in spirals with larger pitch angles.

Many other systems correlate with spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}). Stronger bulges and more massive central black holes have both correlated with more tightly wound spiral arms.

>>>>

Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}, examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}). The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.

@@

A majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and it is possible that they trigger star formation \citep{2013A&A...560A..59C}, though their main role may instead be to sweep up material, promoting the growth of Giant Molecular Clouds. \citep{2014IAUS..298..221D}. This rearrangement of disc gas can lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g. \citealt{2004ARA&A..42..603K}), which are prevalent in most spiral galaxies \citep{2018MNRAS.473.4731K}. Studies of spiral morphology have found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} not \citealt{2019MNRAS.487.1808M} found such a relation in large samples); and spiral tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure). These predictions and observations provide compelling reasons for investigating the underlying rules and dynamics of spiral stucture, as doing so is essential for understanding the secular evolution of disc galaxies.

Our current understanding of the mechanisms which drive spiral growth and evolution suggests that each of the different forms of spiral galaxy may be triggered by different processes. Grand Design spirals are thought to have undergone a tidal interaction \citep{2010MNRAS.403..625D,2017ApJ...834....7S}, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S,2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and recurrent in nature \citep{1966ApJ...146..810J}.

One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanged for many rotational periods \citep{1964ApJ...140..646L}. The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational: most disc galaxies show spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.

Many simulations demonstrate that spirals do not maintain a constant tightness (often quantified by pitch angle, the angle between the spiral and the tangent to a circle centred on the galaxy, \citealt{1987gady.book.....B}, illustrated in Figure \ref{fig:pitch-angle-example}), and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}. Recent research suggests that spirals arms are transient in nature, and continually dissipate and re-form \citep{2014PASA...31...35D}. These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e. ``wave amplification by stimulated emission of radiation'', \citealt{1976ApJ...205..363M}; swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave. The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle, before dissipating to be reformed.

\begin{figure}
  \includegraphics[width=8.4cm]{plots/pitch-angle-explanation.pdf}
  \caption{Illustration of the definition of pitch angle. It is given as $\phi = \tan^{-1}\left(\frac{\mathrm{d}r}{\mathrm{d}\theta}\ /\ r\right)$, or the angle between the spiral (red) and the tangent to a circle centred on the galaxy (blue).}
  \label{fig:example-spiral-fits}
\end{figure}
In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissipation. \citet{2019arXiv190910291P} propose a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time. They found that the distribution of pitch angles of their sample of 86 galaxies was consistent with this prediction, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.

We aim to test this idea of spiral winding using data from the \textit{Galaxy Builder} citizen science project for the spiral galaxies present in \citet{2020arXiv200610450L}. We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arms produced by \textit{Galaxy Builder}. This methodology allows us to quantify the differences in pitch angles between arms in a single galaxy, as well as investigate the distribution of pitch angles in the galaxy population and investigate relationships between pitch angle and galaxy morphology.

Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of a stellar bar, simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}. Manifold theory is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds. One of the primary factors influencing the shape of this invariant manifold is the relative strength of the non-axisymmetric forcing caused by the bar, with stronger bars resulting in spirals with larger pitch angles.

Many other components correlate with spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}). Larger bulges and more massive central black holes have both correlated with more tightly wound spiral arms.

### 2020-06-23 sections/method.tex
makes changes suggested by SK

% !TEX root = ../main.tex
\subsection{Measuring galaxy pitch angle}

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), Fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}). One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} \citep{2020arXiv200610450L}. These methods attempt to separate light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \citet{2020arXiv200610450L} propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.

@@

\begin{equation}
r = A\,e^{\theta\tan\phi},
\end{equation}
%
where $\phi$ is the arm's pitch angle. One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}). Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e. for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$

@@

A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified using \textsc{SpArcFiRe} were identified as ``good'' matches to real spiral arms by citizen science classifiers.

Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles. Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and reports the pitch angle with the highest amplitude as the galaxy's pitch angle. \citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with visual tracing of spiral arms, to great effect. It is uncertain how the variation between pitch angles of individual arms impacts this measurement.

@@

Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency. We combine the classifications of galaxies in this \textit{validation subset} with the original classifications. Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}. We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 139 galaxies, 261 spiral arms and 239,947 points.

@@

Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \citet{2020arXiv200610450L}. Arms are individually corrected to all have the same chirality (a pitch angle greater than or equal to zero) using the logarithmic spiral fit in \citet{2020arXiv200610450L}.

\subsection{Bayesian modelling of spiral arms in \textit{Galaxy Builder}}

In this Section, we lay out our bayesian hierarchical model for galaxy pitch angle. We fit directly to clustered, cleaned points from polylines drawn in \textit{Galaxy Builder}, deprojected and unwrapped to polar coordinates. We fit a logarithmic spiral to each clustered arm, with the pitch angles of multiple arms in a single galaxy being drawn from a single parent distribution.

@@

Where $\widetilde{r_\mathrm{arm}}$ is the model's predictions for the radii of the deprojected points in a \textit{Galaxy Builder} arm ($\widevec{r_\mathrm{arm}}$), and $\widevec{\theta_\mathrm{arm}}$ is the polar angles of the points.

>>>>

% !TEX root = ../main.tex
\subsection{Measuring galaxy pitch angle}
Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), Fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}). One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} \citep{2020arXiv200610450L}. These methods attempt to separate light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \citet{2020arXiv200610450L} propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.

@@

\begin{equation}
  \label{eq:log-spiral}
r = A\,e^{\theta\tan\phi},
\end{equation}
%
where $\phi$ is the arm's pitch angle, $A$ is an amplitude coefficient and $\theta$ is the polar coordinate. One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}). Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e. for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$

@@

A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified using \textsc{SpArcFiRe} \citep{2014ApJ...790...87D} were identified as ``good'' matches to real spiral arms by citizen science classifiers.

Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles. Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and reports the pitch angle with the highest amplitude as the galaxy's pitch angle. \citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with visual tracing of spiral arms, succesfully eliminating observed bias in a sample of toy images of galaxies. It is uncertain how the variation between pitch angles of individual arms impacts this measurement.

@@

Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency. We combine the 30 classifications of galaxies in this \textit{validation subset} with the 30 original classifications. Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}. We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 139 galaxies, 261 spiral arms and 239,947 points.

@@

Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \citet{2020arXiv200610450L}. Arms are individually corrected to all have the same chirality (a pitch angle greater than or equal to zero) using the logarithmic spiral fit in \citet{2020arXiv200610450L}. This was achieved by multiplying the polar coordinate $\theta$ by $-1$ for arms identified as winding counter-clockwise.

\subsection{Bayesian modelling of spiral arms in \textit{Galaxy Builder}}
\label{section:bhsm-model}
In this Section, we lay out our Bayesian hierarchical model for galaxy pitch angle. We fit directly to clustered, cleaned points from polylines drawn in \textit{Galaxy Builder}, deprojected and unwrapped to polar coordinates. We fit a logarithmic spiral to each clustered arm, with the pitch angles of multiple arms in a single galaxy being drawn from a single parent distribution.

@@

Where $\widetilde{r_\mathrm{arm}}$ is the model's predictions for the radii of the deprojected points in a \textit{Galaxy Builder} arm ($\widevec{r_\mathrm{arm}}$), $c_\mathrm{arm}$ is the amplitude parameter (equivalent to $A$ in Equation \ref{eq:log-spiral}), and $\widevec{\theta_\mathrm{arm}}$ is the polar angles of the points.

### 2020-06-23 sections/results.tex
makes changes suggested by SK

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with less than {1.6\degree} of uncertainty for 95\% of arms, assuming no error on disc inclination and position angle. The pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$), however, is not well constrained. This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of ($\sigma_{\phi_\mathrm{gal}}$) of  {7.9\degree}, which decreases to {6.8\degree} and {6.0\degree} for galaxies with three and four arms respectively. This is roughly consistent with the standard error on the mean for a galaxy with $N$ arms,

\begin{equation}
  \sigma_{\phi_\mathrm{gal}} = \frac{\sigma_\mathrm{gal}}{\sqrt{N}},
\end{equation}
where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 0.9^\circ$. This variability is similar to that found by \citet{1981AJ.....86.1847K} and \citet{2014ApJ...790...87D} and emphasises the need for fitting algorithms to not assume all arms have the same pitch angle. The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy. We see a slight drop in $E[\phi_\mathrm{gal}]$ compared to $E[\phi_\mathrm{arm}]$ at small galaxy pitch angles, due to the truncation of $\phi_\mathrm{gal}$ at {0\degree}.

@@

We see no correlation between galaxy pitch angle derived from the hierarchical model and $B_\mathrm{avg}$. The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) and $B_\mathrm{avg}$ is 0.00.

@@

One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}. In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a good measure of the torque applied on the disc gas.

We do not observe a correlation between $p_\mathrm{bar}$ and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05). Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($p_\mathrm{bar} < 0.2$), with a weak bar ($0.2 \le p_\mathrm{bar} \le 0.5$) and with a strong bar ($p_\mathrm{bar} > 0.5$). Performing marginalized three-sample Anderson-Darling tests does not find that pitch angles ($\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$) of galaxies with different bar strengths were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples for the test of $\phi_\mathrm{gal}$, and at the 10\% level for the test of $\phi_\mathrm{arm}$. The distributions of Anderson-Darling test statistic is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.

@@

We do not correct for variation in bulge size in this work, however, predictions from Manifold theory should not be affected as bulges do not provide a strong non-axisymmetric forcing. The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.

@@

We aim to replicate their work here, using our sample and methods. We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis. Observation of the distribution of arm pitch angles in our sample suggests limits of $15^\circ < \phi < 50.0^\circ$.

@@

This result suggests that we cannot rule out a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements. This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples. As we have no information available on the biases present in \textit{Galaxy Builder} spiral arm classification, we choose to keep the less strict limit of {15\degree}.

\subsubsection{Arm Pitch angle}
The inconclusive result for $\phi_\mathrm{gal}$ is perhaps unsurprising: were we to assume that spiral arms are transient and recurrent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time. This is supported by the large observed spread in inter-arm pitch angles.

If we assume instead that spirals form and wind independently in a galaxy, and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of the cotangent of pitch angles of individual arms should be uniform between our limits, rather than that of the galaxy's pitch angle as a whole.

>>>>

% !TEX root = ../main.tex
\subsection{Constraints on Galaxy Pitch angle}
\label{section:constraints-on-galaxy-phi}
Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with posterior standard deviations less than {1.6\degree} for 95\% of arms, assuming no error on disc inclination and position angle. This is illustrated well by the small uncertainties on fit spiral arms in Figure \ref{fig:example-spiral-fits}. The pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$), however, is not well constrained. This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles. For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of ($\sigma_{\phi_\mathrm{gal}}$) of  {7.9\degree}, which decreases to {6.8\degree} and {6.0\degree} for galaxies with three and four arms respectively. This is roughly consistent with the standard error on the mean for a galaxy with $N$ arms,

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/example-spiral-fits.pdf}
  \caption{Examples of spiral profiles fit using the hierarchical model described in Section \ref{section:bhsm-model}. Deprojected points from \textit{Galaxy Builder} clustered, cleaned spiral arms are shown in black; fit logarithmic spiral arms are shown in red, with the width of the line corresponding to the $2\sigma$ interval on predicted values of $\widetilde{r_\mathrm{arm}}$.}
  \label{fig:example-spiral-fits}
\end{figure*}

\begin{equation}
  \sigma_{\phi_\mathrm{gal}} = \frac{\sigma_\mathrm{gal}}{\sqrt{N}},
\end{equation}
where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 0.9^\circ$. This variability is similar to that found by \citet{1981AJ.....86.1847K} and \citet{2014ApJ...790...87D} and emphasises the need for fitting algorithms to not assume all arms have the same pitch angle. The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy. We see a slight drop in the expectation values of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) compared to the expectation of arm pitch angles ($E[\phi_\mathrm{arm}]$) at small galaxy pitch angles, due to the truncation of $\phi_\mathrm{gal}$ at {0\degree}.

@@

We see no correlation between galaxy pitch angle derived from the hierarchical model and $B_\mathrm{avg}$. The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) and $B_\mathrm{avg}$ is 0.00 (with a p-value of 0.95).

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/bulge_bar_phigal_distribution.pdf}
  \caption{Density plot showing bulge strength ($B_\mathrm{avg}$; left, orange) and bar strength ($p_\mathrm{bar}$; right, green) against galaxy pitch angle ($\phi_\mathrm{gal}$). Split points for the marginalized Anderson-Darling tests are labelled. There is no statistically significant relationship for either bulge or bar strength.}
  \label{fig:bulge-bar-pa-hist}
\end{figure*}

@@

One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}. In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.

We do not observe a correlation between $p_\mathrm{bar}$ and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05, with a p-value of 0.54). Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($p_\mathrm{bar} < 0.2$), with a weak bar ($0.2 \le p_\mathrm{bar} \le 0.5$) and with a strong bar ($p_\mathrm{bar} > 0.5$). Performing marginalized three-sample Anderson-Darling tests does not find that pitch angles ($\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$) of galaxies with different bar strengths were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples for the test of $\phi_\mathrm{gal}$, and at the 10\% level for the test of $\phi_\mathrm{arm}$. The distributions of Anderson-Darling test statistic is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.

@@

The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.

@@

We perform a similar test in this work, using our sample and methods. We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis. Observation of the distribution of arm pitch angles in our sample (Figure \ref{fig:pa-cot-distributions}) suggests limits of $15^\circ < \phi < 50.0^\circ$.

@@

This result suggests that we cannot rule out a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements. This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples. As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.

\subsubsection{Arm Pitch angle}
The inconclusive result for $\phi_\mathrm{gal}$ is perhaps unsurprising: were we to assume that spiral arms are transient and recurrent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time. This is supported by the large observed spread in inter-arm pitch angles (Section \ref{section:constraints-on-galaxy-phi}).

If we assume instead that spirals form and wind independently inside a galaxy, and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of the cotangent of pitch angles of individual arms should be uniform between our limits, rather than that of the galaxy's pitch angle as a whole.

### 2020-06-23 sections/summary.tex
makes changes suggested by SK

The methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy. As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed. A larger sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.

>>>>

The methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy. As with most analyses, the most impactful improvement it would be possible to make here would be to increase the quality and volume of data analysed. A larger sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.
